---
title: "FinalProject"
author: "Group 3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tigerstats)
library(readr)
library(tidyverse)
library(webr)
```

## Section 1 - Loading the data and Exploring the data

We have selected Bike dataset to perform exploration of the data. We start by loading the data into the R Markdown file

```{r}
bike_data <- read_csv("Cleaned_bike_data.csv",show_col_types = FALSE)
```


```{r}
summary(bike_data)
```

We can see that there are different columns in the table which are Model_name, Model_year, Kms_driven, Owner, Location, Mileage, Power and Price.

To determine the different types of the datatypes, we can use the str command in R. This command tells us the different types of data which is present in the datasheet.
```{r}
str(bike_data)
```

The different types of datatypes which are present in the dataset are numeric and charachter.  

To determine the null values in the table, we can use the built in method in R. 

```{r}
null_values <- sum(is.na(bike_data)) 
null_values
```

With the above method, we can see that there are 0 missing/Null values in the dataset.

## Part 2 - Graphical Overview

You can also embed plots, for example:

```{r pressure, echo=FALSE}
```

## Part 3 - Hypothesis Testing

```{r }
data <- read.csv("Cleaned_bike_data.csv")
str(data)
```

### Testing First hand owner price higher the Second hand owner (One-side test)

\begin{align*}
H_0: \mu_{first\_owner\_price}-\mu_{second\_owner\_price} & = 0 \\
H_1: \mu_{first\_owner\_price}-\mu_{second\_owner\_price} & > 0 
\end{align*}

```{r , fig.width=8, fig.height=7}


#split first hand owner and second hand owner
first_hand_owner <- subset(data, owner=="first owner")
second_hand_owner <- subset(data, owner=="second owner")


first_second_owner <- rbind(first_hand_owner, second_hand_owner)


ggplot(first_second_owner, aes(x=owner, y=price, fill=owner)) + 
    geom_boxplot(alpha=0.8) +
    theme(legend.position="none") +
    scale_fill_brewer(palette="Dark2") +
    coord_flip()
```
#### Don't know variance. So, use T-testing but need to check variance whether or not equal. \
#### Check variance equal or not \

\begin{align*}
H_0: \sigma_{first\_owner\_price}^2 = \sigma_{second\_owner\_price}^2 \\
H_1:  \sigma_{first\_owner\_price}^2 \ne \sigma_{second\_owner\_price}^2 
\end{align*}

```{r , fig.width=8, fig.height=7}
var.test(first_hand_owner$price, second_hand_owner$price, alternative = "two.sided")
#Not reject H0 => variance equal

#t-testing
t.test(first_hand_owner$price, second_hand_owner$price,alternative="greater", var.equal = TRUE, conf.level = 0.95)

plot(t.test(first_hand_owner$price, second_hand_owner$price,alternative="greater", var.equal = TRUE, conf.level = 0.95))

```
\
>Should not reject Mu(first_price)=Mu(second_price) 


### Testing Price of Second hand bike not be different from third and later (Two-side test) \

\begin{align*}
H_0: \mu_{second\_owner\_price}-\mu_{later\_second\_owner\_price} & = 0 \\
H_1: \mu_{second\_owner\_price}-\mu_{later\_second\_owner\_price} & > 0 
\end{align*}


```{r, fig.width=8, fig.height=7}
second_hand_owner <- subset(data, owner=="second owner")
later_second_hand_owner <- subset(data, owner =="third owner" | owner == "fourth owner or more")

later_second_hand_owner$owner[later_second_hand_owner$owner == "third owner" | later_second_hand_owner$owner == "fourth owner or more"]<- "third owner or more"

second_later_owner <- rbind(second_hand_owner, later_second_hand_owner)


ggplot(second_later_owner, aes(x=owner, y=price, fill=owner)) + 
    geom_boxplot(alpha=0.8) +
    theme(legend.position="none") +
    scale_fill_brewer(palette="Dark2") +
    coord_flip()
```


#### Don't know variance. So, use T-testing but need to check variance whether or not equal. \
#### Check variance equal or not \
\begin{align*}
H_0: \sigma_{second\_owner\_price}^2 = \sigma_{later\_second\_owner\_price}^2 \\
H_1:  \sigma_{second\_owner\_price}^2 \ne \sigma_{later\_second\_owner\_price}^2 
\end{align*}

```{r, fig.width=8, fig.height=7}
var.test(second_hand_owner$price, later_second_hand_owner$price, alternative = "two.sided")
#reject H0 => variance not equal

#t-testing
t.test(second_hand_owner$price, later_second_hand_owner$price,alternative="two.sided", var.equal = FALSE, conf.level = 0.95)

plot(t.test(second_hand_owner$price, later_second_hand_owner$price,alternative="two.sided", var.equal = FALSE, conf.level = 0.95))
```
\
> Should reject Mu(second_price)=Mu(later_second_price) => (Mu(second_price) != Mu(later_second_price)


### Testing Higher median kms driven price cheaper than lower median kms driven price (One side test)

\begin{align*}
H_0: \mu_{greater\_median\_kms\_driven}-\mu_{less\_median\_kms\_driven} & = 0 \\
H_1: \mu_{greater\_median\_kms\_driven}-\mu_{less\_median\_kms\_driven} & > 0
\end{align*}

```{r,  fig.width=8, fig.height=7}
median(data$kms_driven)


greater_median_kms_driven <- subset(data, kms_driven > 18000)
less_median_kms_driven <- subset(data, kms_driven <= 18000)

greater_median_kms_driven["median_kms_driven"] <- "greater"
less_median_kms_driven["median_kms_driven"] <- "less"

all_kms_driven <- rbind(greater_median_kms_driven, less_median_kms_driven)


ggplot(all_kms_driven, aes(x=median_kms_driven, y=price, fill=median_kms_driven)) + 
    geom_boxplot(alpha=0.8) +
    theme(legend.position="none") +
    scale_fill_brewer(palette="Dark2") +
    coord_flip()

```

#### Don't know variance. So, use T-testing but need to check variance whether or not equal.
#### Check variance equal or not
\begin{align*}
H_0: \sigma_{greater\_median\_kms\_driven\_price}^2 = \sigma_{less\_median\_kms\_driven\_price}^2 \\
H_1:  \sigma_{greater\_median\_kms\_driven\_price}^2 \ne \sigma_{less\_median\_kms\_driven\_price}^2 
\end{align*}

```{r, fig.width=8, fig.height=7}
var.test(greater_median_kms_driven$price, less_median_kms_driven$price , alternative = "two.sided")
# reject H0 => variance not equal

#t-testing
t.test(greater_median_kms_driven$price, less_median_kms_driven$price,alternative="less", var.equal = FALSE, conf.level = 0.95)

plot(t.test(greater_median_kms_driven$price, less_median_kms_driven$price,alternative="less", var.equal = FALSE, conf.level = 0.95))
```
\
> Should reject H0 Mu(over_mean_kms_driven_price)=Mu(less_mean_kms_driven_price) => Mu(over_mean_kms_driven_price)< Mu(less_mean_kms_driven_price)




## Part 4 - Linear Regression

This section is for Linear Regression.
